Investigating social change during cultural contact period using geometric morphometry of pottery shapes from Iron Age northeastern Taiwan

### 2020-01-31 analysis/paper/paper.Rmd
paste the deleted para about metric data to scraps.R

This study demonstrates an application of elliptic Fourier analysis on ceramic shapes to explore the emergence of ceramic specialization as indicative of foreign influences. Here, EFA is combined with a significance test for the equality of CVs of shape variables to provide a robust way to identify and statistically assess differences in shape standardization. The direct relationship between foreign influences and standardization of ceramic shape was tested on pottery from Kiwulan, and large Iron Age Indienous settlement in northeast Taiwan. Lower variation in ceramic shape was identified after European presence began, and even lower variation during the period of Chinese presence. Our findings help to understand the factors that may lead to standardization of pottery production in a cross-cultural interaction context. More homogeneous shapes during the contact periods, without any changes in clay paste composition or production technique, suggests that shape standardization was intentional. The results here further suggest that expressions of social identity or cultural boundaries in Indigenous societies through highly visible vessel qualities, such as shape, may be heightened during periods of foreign intrusion in pericolonial contexts. Our analysis, with its openly available methods and data, is readily extensible to other pottery assemblages in the region to further explore related questions about craft specialization and standardization in Iron Age ceramic technologies.

>>>>

This study demonstrates an application of elliptic Fourier analysis on ceramic shapes to explore the emergence of ceramic specialization as indicative of foreign influences. Here, EFA is combined with a significance test for the equality of CVs of shape variables to provide a robust way to identify and statistically assess differences in shape standardization. The direct relationship between foreign influences and standardization of ceramic shape was tested on pottery from Kiwulan, and large Iron Age Indigenous settlement in northeast Taiwan. Lower variation in ceramic shape was identified after European presence began, and even lower variation during the period of Chinese presence. Our findings help to understand the factors that may lead to standardization of pottery production in a cross-cultural interaction context. More homogeneous shapes during the contact periods, without any changes in clay paste composition or production technique, suggests that shape standardization was intentional. The results here further suggest that expressions of social identity or cultural boundaries in Indigenous societies through highly visible vessel qualities, such as shape, may be heightened during periods of foreign intrusion in pericolonial contexts. Our analysis, with its openly available methods and data, is readily extensible to other pottery assemblages in the region to further explore related questions about craft specialization and standardization in Iron Age ceramic technologies.

### 2020-01-31 analysis/paper/scraps.Rmd
delete one script file
paste the deleted para about metric data to scraps.R

# Part of Introdution

@@

>>>>

# Part of Introduction

@@

#--------------original metric data analysis------------
Table \@ref(tab:metricCVstest) shows the results of Modified signed-likelihood ratio test (MSLRT) for equality of CVs for for six measurements and two ratio values. There is a significant difference between CV values for two metric measurements across phases, the rim diameter and the neck diameter. The CV values show that the variation of rim diameter is small (6.8%)  before European contact, increases after the European presence (9.6%), and decreases after the Chinese presence (6.9%). For the neck diameter, the CV values indicate smaller variation after the European period (7.7%) and the Chinese period (7.1%) compared to pre-European period (10.6%). <!-- these numeric values in this paragraph should come from inline code, not copy-pasted in like this -->
<!-- You need to put the code that computes the MSLRT results right in here, in the Rmd -->

```{r metricCVstest}
metric_cvs_with_phase <- readr::read_csv(here::here("analysis", "data", "raw_data", "Kwl_p_metrics_cv_test.csv"))
metric_cvs_with_phase_tidy <-
  metric_cvs_with_phase %>% 
  mutate_at(c("pre-e", "post-e", "ch-con", "MSLRT", "p_value"), round, 4) %>% 
  rename(`pre-European` = "pre-e", `after European` = "post-e", Chinese = "ch-con")
knitr::kable(metric_cvs_with_phase_tidy,
             caption = "Coefficient of Variation test for metric attributes by phases")
```

#------------revised metric data analysis---------------

Table \@ref(tab:metricCVstest-pre-post) and \@ref(tab:metricCVstest-post-chi) shows the results of Asymptotic test [@Feltz1996] for equality of CVs for for six measurements and two ratio values. There is a significant difference (p-value less than 0.01) between CV values for two metric measurements across phases, the rim diameter and the neck diameter before and after the European presnece. The CV values show that the variation of rim diameter is small (6.8%) before European contact, increases after the European presence (9.6%), and decreases after the Chinese presence (6.9%). For the neck diameter, the CV values indicate smaller variation after the European period (7.7%) and the Chinese period (7.1%) compared to pre-European period (10.6%). After the arrival of the Chinese, there is only one metric attribute, the rim diameter, showing significant difference from post-European period (p-value less than 0.05).

```{r metricCVstest-pre-post, eval=FALSE}
# read metric data
kwl_p_ave <- readr::read_csv(here::here("analysis", "data", "raw_data",
                                 "KWL-pottery-metrics-tidy-ave.csv"))
library(cvequality)
# select metric attributes and tidy data for computing CV
kwl_p_metrics_long <-
  kwl_p_ave %>%
  mutate(dia_rim_body = Dia_Rim_ave/Dia_Body_ave,
         thi_rim_body = Thick_Rim_ave/Thick_Body_ave) %>%
  select(period,
         Dia_Rim_ave,
         Dia_Neck_ave,
         Dia_Body_ave,
         Thick_Rim_ave,
         Thick_Neck_ave,
         Thick_Body_ave,
         dia_rim_body,
         thi_rim_body,
         # H_Rim
  ) %>%
  rename(`Rim diameter (mm)` = Dia_Rim_ave,
         `Neck diameter (mm)` = Dia_Neck_ave,
         `Body diameter (mm)` = Dia_Body_ave,
         `Rim thickness (mm)` = Thick_Rim_ave,
         `Neck thickness (mm)` = Thick_Neck_ave,
         `Body thickness (mm)` = Thick_Body_ave,
         `Ratio of Rim/Body diameter` = dia_rim_body,
         `Ratio of Rim/Body thickness` = thi_rim_body
  ) %>%
  gather(variable, value, -period) %>%
  filter(!is.na(period),
         !is.na(value))
# filter pre-e and post-e periods
kwl_p_metrics_long_pre_post <-
  kwl_p_metrics_long %>%
  filter(!period == "ch-con")
# filter post-e and chi periods
kwl_p_metrics_long_post_chi <-
  kwl_p_metrics_long %>%
  filter(!period == "pre-e")
#------------------------------------------CV test for pre-e and post-e pair
kiwulan_pottery_metrics_cvs_plot_cv_test_asym_pre_post <-
  kwl_p_metrics_long_pre_post %>%
  nest(-variable) %>%  # or nest(period, value) %>%
  mutate(asymptotic_test = map(data, # change from mslr_test to asymptotic_test
                               ~bind_cols(asymptotic_test (
                                 .x$value,
                                 .x$period)))) %>%
  unnest(asymptotic_test)
# compute CV
kwl_p_cvs_pre_post <-
  kwl_p_metrics_long_pre_post %>%
  group_by(period, variable) %>%
  summarise(cvs = raster::cv(value, na.rm = TRUE))
# join p-values to CVs
kwl_p_cvs_and_pvalues_pre_post <-
  kwl_p_cvs_pre_post %>%
  left_join(kiwulan_pottery_metrics_cvs_plot_cv_test_asym_pre_post) %>%
  mutate(significant = ifelse(p_value <= 0.05, 'yes', 'no')) %>%
  mutate(variable = factor(variable, ordered = TRUE)) %>%
  arrange(variable)
# making table for CV
kwl_p_cvs_table_pre_post <-
  kwl_p_cvs_and_pvalues_pre_post %>%
  select(-data, -significant) %>%
  pivot_wider(names_from = period, values_from = cvs) %>%
  mutate(variable = factor(variable, levels = c("Rim thickness (mm)",
                                                "Neck thickness (mm)",
                                                "Body thickness (mm)",
                                                "Ratio of Rim/Body thickness",
                                                "Rim diameter (mm)",
                                                "Neck diameter (mm)",
                                                "Body diameter (mm)",
                                                "Ratio of Rim/Body diameter"))) %>%
  arrange(variable) %>%
  select(variable, `pre-e`, `post-e`, `D_AD`, `p_value`) %>%
  mutate_at(c("pre-e", "post-e", "D_AD", "p_value"), round, 4)
knitr::kable(kwl_p_cvs_table_pre_post,
             caption = "Coefficient of Variation test for metric attributes before and after European presence")
```
knitr::kable(metric_cvs_with_phase_tidy,
             caption = "Coefficient of Variation test for metric attributes by phases")
```{r metricCVstest-post-chi, eval=FALSE}
#-----------------------------------CV test for post-e and ch-con pair
kiwulan_pottery_metrics_cvs_plot_cv_test_asym_post_chi <-
  kwl_p_metrics_long_post_chi %>%
  nest(-variable) %>%  # or nest(period, value) %>%
  mutate(asymptotic_test = map(data, # change from mslr_test to asymptotic_test
                               ~bind_cols(asymptotic_test (
                                 .x$value,
                                 .x$period)))) %>%
  unnest(asymptotic_test)
# compute CV
kwl_p_cvs_pre_post <-
  kwl_p_metrics_long_post_chi %>%
  group_by(period, variable) %>%
  summarise(cvs = raster::cv(value, na.rm = TRUE))
# join p-values to CVs
kwl_p_cvs_and_pvalues_post_chi <-
  kwl_p_cvs_pre_post %>%
  left_join(kiwulan_pottery_metrics_cvs_plot_cv_test_asym_post_chi) %>%
  mutate(significant = ifelse(p_value <= 0.05, 'yes', 'no')) %>%
  mutate(variable = factor(variable, ordered = TRUE)) %>%
  arrange(variable)
# making table for CV
kwl_p_cvs_table_post_chi <-
  kwl_p_cvs_and_pvalues_post_chi %>%
  select(-data, -significant) %>%
  pivot_wider(names_from = period, values_from = cvs) %>%
  mutate(variable = factor(variable, levels = c("Rim thickness (mm)",
                                                "Neck thickness (mm)",
                                                "Body thickness (mm)",
                                                "Ratio of Rim/Body thickness",
                                                "Rim diameter (mm)",
                                                "Neck diameter (mm)",
                                                "Body diameter (mm)",
                                                "Ratio of Rim/Body diameter"))) %>%
  arrange(variable) %>%
  select(variable, `post-e`, `ch-con`, `D_AD`, `p_value`) %>%
  mutate_at(c("post-e", "ch-con", "D_AD", "p_value"), round, 4)
knitr::kable(kwl_p_cvs_table_post_chi,
             caption = "Coefficient of Variation test for metric attributes between European and Chinese presence")
